---
title: "FML Final Exam"
author: "Shiva Chaitanya Goud Gadila"
date: "2023-05-03"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#IMPORTING OF DATASET 
```{r}
PUDL<- read.csv("~/Downloads/fuel_receipts_costs_eia923.csv")

#importing the dataset PUDL(Fuel receipts)
```
#IMPORTING THE LIBRARY 
```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(caret)

#Recalling the istalled packages.
```

#CHECKING OF THE DATA 
```{r}
str(PUDL) 
```
#EXPLORING THE GIVEN DATA
```{r}
glimpse(PUDL)
#Investigating The given Information.
```
#CLEANING THE GIVEN DATA 
While the data itself is clean, the dataset contains several variables that have significant missing values.
Follows these steps:
```{r}

#1. Identify all variables that have significant missing values.
# col names with missing values
colnames(PUDL)[colSums(is.na(PUDL)) > 0]

# all missing values
all <- PUDL %>% 
  summarise_all(funs(sum(is.na(.)))) %>% 
  gather(key = "variable", value = "missing_values") %>% 
  filter(missing_values > 0) %>% 
  arrange(desc(missing_values))

# Remove variables with significant missing values
PUDL <- PUDL %>% 
  select(-all$variable)
```
# CHECKING THE DATA
```{r}
 str(PUDL)
```
2.	Ensure that the variables have the right attributes. For example, numerical or categorical.
```{r}
# attributes
sapply(PUDL, class)
#It determines the data types of each variable in the "PUDL" dataset and returns the information as a vector using the class() function and sapply().
```
3.	To ensure that both the data, and the analysis are unique to each student, randomly sample about 2% of your data using a random 4-digit number as the seed to sample the data. Use 75% of the sampled data as the training set, and the rest as the test set (if needed). This should yield a training set of about 9000 and a test of about 3000. 
```{r}
# Set a random seed for reproducibility
set.seed(1234)

# Randomly sample about 2% of the rows from the dataset
sampled <- PUDL %>% 
  sample_frac(0.02)

# Split the sampled data into training and test sets

train<- sampled %>% 
  sample_frac(0.75)

test<- sampled %>%
    anti_join(train)

# Print the number of rows in the training and test sets
nrow(train)
nrow(test)

#This code randomly samples 2% of the "PUDL" dataset, splits it into training and test sets, and prints the number of rows in each set. It is a common data preparation step before building and evaluating predictive models.
```
#CHECKING THE DATA 
```{r}
str(train)
str(test)
```
```{r}
# Identify the numeric variables in the training set
numVars <- names(train)[sapply(train, is.numeric)]

# Create the scatterplot matrix using the pairs() function
pairs(train[,numVars], pch = 19, cex = 0.5,) 

#The code selects the numeric variables from the train dataframe and stores them in numVars. It then creates a scatterplot matrix using the pairs() function to visualize the pairwise relationships between the numeric variables in the train dataset.
      
```

```{r}
# k-means clustering
set.seed(1234)
# Select numeric columns from the train data
numValues <- train %>% select_if(is.numeric)

# Perform k-means clustering with 3 clusters
kmeans <- kmeans(numValues, centers = 3)
kmeans

#It performs k-means clustering with 3 clusters on the numeric columns of the training data. The resulting clusters will be stored in the kmeans object.
```
```{r}
# aggregate the data
aggregate(train[,numVars], by = list (kmeans $ cluster), mean)

#This code aggregates the numeric variables in the training set by the k-means cluster labels and computes the mean for each variable within each cluster.
```

#visualizing the given data
```{r}
#  Create a ggplot object with the train data and cluster assignments on the y-axis
ggplot(train, aes(y = kmeans$cluster)) +

# Add a bar layer to the plot, with fill color based on cluster assignments
  geom_bar(aes(fill = kmeans$cluster), position = "dodge") + 

#  Apply the economist theme to the plot
  theme_economist() + 

# Adjust the title alignment
  theme(plot.title = element_text(hjust = 0.5))

#This creates a bar plot using ggplot, with the cluster assignments on the y-axis and the fill color based on the cluster assignments. The economist theme is applied to the plot, and the title alignment is adjusted.
```
#KNN
```{r}
# Set the seed for reproducibility
set.seed(1234)

# Identify numeric columns in the train data
numValues <- sapply(train, is.numeric)

# Load the 'class' package for kNN classification
library(class)

#Subset the train and test data to only include numeric columns
train1 <- train[, numValues]
test1 <- test[, numValues]

# Perform kNN classification on the numeric train and test data
# using the cluster assignments from the k-means model as the class labels
knn <- knn(train1, test1, cl = kmeans$cluster, k = 3)
knn

#It perform k-Nearest Neighbors classification on the numeric columns of the train and test data, using the cluster assignments obtained from k-means clustering with 3 clusters on the train data as the class labels. The resulting classification is stored in the knn object.
```
#SEGMENTATION 
```{r}
#Identify numeric columns in the train data
numValues <- sapply(train, is.numeric)

# Step 2: Perform k-means clustering on the numeric train data, with 3 clusters
kmeans <- kmeans(train[, numValues], centers = 3)

kmeans

#This code performs k-means clustering on the numeric columns of the training data with 3 clusters. The resulting object is saved as kmeans.
```

```{r}
# Step 1: Get the length of the k-means cluster assignments
kmeans_length <- length(kmeans$cluster)
kmeans_length 

# Step 2: Get the length of the kNN classifications
knn_length <- length(knn)
knn_length

# It calculates the length of the k-means cluster assignments and the length of the kNN classifications, stored in the variables kmeans_length and knn_length, respectively.
```




